---
title: Bonanza creek snowshoe hare counts and sizes - exploratory report (CORNELIUS)
author: "Andrea Cornelius"
date: "11/3/2021"
output: html_document

---
### 1. Overview

This report provides an exploratory overview of differences in trapping counts and weights for juvenile snowshoe hares (Lepus americanus) trapped in the Bonanza Creek Experimental Forest, (a U.S. Long Term Ecological Research Network site) in Fairbanks, Alaska, over fourteen years, from 1999 through 2012 (Kielland et al 2017). Hares were caught in several traps within each of four different sample grids, identified through ear tags, and were weighed, measured, aged, and sexed. Specifically, juvenile hares trapped are counted, weights are analyzed by site and sex, and the relationship(s) between weight and hindfoot length is explored.

### 2. Data and methods

Snowshoe hare structural size measurements were collected and made available by Dr. Knut Kielland and colleagues (2017) at the Bonanza Creek Long Term Ecological Research (LTER) network site in Fairbanks, Alaska. The data contain observations for 3197 hare trappings, collected over thirteen years, from 1999-2012. Following exploratory data visualization, weights between male and female juvenile hares, and at different trapping grids, are analyzed by unpaired t-tests using a significance level ($\alpha$) of 0.05 throughout. Differences between groups are described by Cohen's *d* effect size. The relationship between weight and hindfoot length is explored by simple linear regression. All analyses are in R version 5.12.10 using RStudio version 1.4.1717.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
# Attach packages
library(tidyverse)
library(janitor)
library(here)
library(lubridate)
library(kableExtra)
library(effsize)
library(ggpubr)
library(broom) 
library(equatiomatic)
library(ggbeeswarm)
```


```{r}
# Read in data
hares <- read_csv(here("data", "bonanza_hares.csv"))
```

```{r, include = FALSE}
# Date classification
date <- as.Date("1998-11-26") # Change date class to date

date

class(date) # Check the class
```

```{r, include = FALSE}
# Weight classification 
weight <- as.double() # Classify weight class to double

weight # Call it out

class(weight) # Check the class
```

### 3. Juvenile hare initial investigation

#### A. Annual juvenile hare trap comparison and descriptive statistics

```{r}
# Count the total number of juvenile hare trappings during each year of the study
juvenile_traps <- hares %>% # Pull from hares data
  mutate(date_new = mdy(date)) %>% # Convert class to date
  mutate(year = year(date_new)) %>% # Add new column for year
  filter(age == "j") %>% # Filter age to juvenile
  group_by(year) %>% # Group by year
  summarize("count" = n()) # Get counts
```

Table 1. Descriptive statistics (mean, median, standard deviation, minimum and maximum count) for juvenile hares trapped at Bonanza Creek LTER from 1999-2012. Data: Kielland et al. 2017.
```{r, descriptive_stats}
# Get stats for trapped juveniles
juvenile_traps_stats <- juvenile_traps %>%
  summarize("Mean Count" = round(mean(count), 0),
            "Median Count" = round(median(count), 0),
            "Standard Deviation Count" = round(sd(count), 0),
            "Minimum Count" = round(min(count), 0),
            "Maximum Count" = round(max(count), 0))

# Make a finalized Kable table
j_traps_table <- knitr::kable(head(juvenile_traps_stats)) %>%
  kable_styling(full_width = F)
```

```{r}
# Data visualization of the total number of juvenile hare trappings during each year of the study
ggplot(data = juvenile_traps, aes(x = year, y = count)) +
  geom_bar(stat = "identity", fill="#f68060") +
  scale_x_continuous(breaks = 1999:2012) +
  scale_y_continuous(limits=c(0, 130), breaks = c(0, 25, 50, 75, 100, 125)) +
  theme_pubclean() +
   labs(x = "Year",
       y = "Juvenile Snowshoe Hare Trappings") +
  theme(text = element_text(family = "Times New Roman"))
```

**Figure 1.** Count of juvenile snowshoe hares trapped in the Bonanza Creek LTER from 1999-2012.

Juvenile snowshoe hare trappings were compared across fourteen years, from 1999-2012; sex and grid are not included as variables here, but should be investigated moving forward. To standardize the juvenile hare population in future analyses, frequency of checking traps and number of traps across grids should more or less correlate year to year. In the data here, the minimum trapping count was `r juvenile_traps_stats[1,4]` juvenile hares and the maximum trapping count was `r juvenile_traps_stats[1,5]` juvenile hares. The mean juvenile hare count was (`r juvenile_traps_stats[1,1]` $\pm$ `r juvenile_traps_stats[1,3]`; mean $\pm$ 1 standard deviation) and the median juvenile hare count was `r juvenile_traps_stats[1,2]`. (Table 1, Figure 1).

#### B. Difference in weight betweeen juvenile hares by sex and site

```{r}
# Data wrangling to count the total number of juvenile hare trappings by grid and sex
hare_subset <- hares %>% # Pull from hares data
  filter(age == "j") %>% # Filter age to juvenile
  group_by(grid, sex) %>% # Group by grid and sex
  select(-date, -time, -trap, -l_ear, -r_ear, -hindft, -notes, -b_key, -session_id, -study) %>%
  mutate(sex = case_when( # Change cell abbreviations to full names
         sex == "f" ~ "Female",
         sex == "m" ~ "Male",
         is.na(sex) == TRUE ~ "Unknown"),
         grid = case_when(
         grid == "bonrip" ~ "Bonanza Riparian",
         grid == "bonmat" ~ "Bonanza Mature",
         grid == "bonbs" ~ "Bonanza Black Spruce"))
``` 


```{r}
# Data visualization of the total number of juvenile hare trappings by grid and sex
ggplot(data = hare_subset, aes(x = sex, y = weight, color = sex)) +
  geom_jitter(size = 2) +
  scale_color_manual(values = c("pink", "blue", "grey"),
                     name = "Grids") +
  facet_wrap(~ grid) +
  labs(x = "Sex",
       y = "Weight (g)") +
  theme_cleveland() +
  theme(legend.position = "none",
        axis.title.x = element_text(size = 12),
        axis.title.y = element_text(size = 12),
        strip.text = element_text(size = 12, face = "bold"))
```

**Figure 2.** Juvenile snowshoe hare weight distribution by grid and sex.

On average, male juvenile snowshoe hares had a greater weight than female juvenile snowshoe hares across grids. Both male and female hares trapped at Bonanza Black Spruce had higher mean weights than hares of their respective sexes trapped at other grids, though there were the fewest samples there, so that may not be accurate if there were more samples.

```{r, include = FALSE}
# Get stats for juvenile hares by sex and grid
weight_summary_1 <- hare_subset %>%
  group_by(sex, grid) %>%
  summarize("Mean Weight (g)" = mean(weight, na.rm = TRUE),
            "Median Weight (g)" = median(weight, na.rm = TRUE),
            "Standard Deviation (g)" = sd(weight, na.rm = TRUE),
            "Sample Size" = n()) %>%
  rename("Sex" = sex, "Grid" = grid)

# Finalize table with Kable
knitr::kable(head(weight_summary_1),
  caption = "<b> Table 2: Mean, Median, and Standard Deviations for juvenile snowshoe hares by sex. <b>") %>%
  kable_styling(full_width = F)
```

#### C. Difference in weight between male and female juvenile snowshoe hares

**Table 2.** Mean, Median, and Standard Deviations for juvenile snowshoe hares by sex.

```{r}
# Get stats for juvenile hare weight by sex
sex_weight <- hare_subset %>%
  group_by(sex) %>%
  summarize("Mean Weight (g)" = round(mean(weight, na.rm = TRUE), 2), 
            "Median Weight (g)" = round(median(weight, na.rm = TRUE), 2),
            "Standard Deviation (g)" = round(sd(weight, na.rm = TRUE), 2),
            "Sample Size" = n(), na.rm = TRUE) %>%
  rename("Sex" = sex) %>%
  select(-na.rm)

# Finalize summary table through Kable
knitr::kable(head(sex_weight)) %>%
  kable_styling(full_width = F)
```

```{r, include = FALSE}
# Make subsets of female and male juvenile weights to do unpaired t test and Cohen's d
female_weight <- hare_subset %>%
  filter(sex == "Female") %>%
  drop_na(weight) %>%
  pull(weight)

male_weight <- hare_subset %>%
  filter(sex == "Male") %>%
  drop_na(weight) %>%
  pull(weight)

# Run the t.test
weight_ttest <- t.test(male_weight, female_weight, na.rm = TRUE)

# Get the tidy model results to call outputs in-line: 
weight_ttest_tidy <- tidy(weight_ttest)

# Get the means & SDs for each: 
male_mean <- mean(male_weight)
female_mean <- mean(female_weight)
male_sd <- sd(male_weight)
female_sd <- sd(female_weight)

# Get the effect size: 
weight_effsize <- (cohen.d(male_weight, female_weight))

weight_ttest

weight_effsize
```

On average, juvenile snowshoe hare males have a greater weight than juvenile snowshoe hare females (`r (male_mean)` $\pm$ `r (male_sd)` and `r (female_mean)` $\pm$ `r (female_sd)` g, respectively; mean $\pm$ 1 standard deviation). While the absolute difference in means is `r (male_mean) - (female_mean)` g (a `r ((male_mean - female_mean)/((male_mean + female_mean)/2)*100)`% difference), the difference in means is significant (Welch's two-sample t-test: t(`r (weight_ttest_tidy$parameter)`) = `r (weight_ttest_tidy$statistic)`, p = 0.007), and the effect size is small (Cohen's *d* = `r (weight_effsize$estimate)`).

```{r, include = FALSE}
male_female <- hare_subset %>% 
  filter(sex %in% c("Male", "Female"))
```

```{r, include = FALSE}
# Make exploratory histogram to compare sex effect on weight
ggplot(data = male_female, aes(x = weight)) +
  geom_histogram(bins = 15) +
  facet_wrap(~ sex) +
  scale_fill_manual(values=c("pink", "blue"),
                    name = "sex") +
  labs(x = "Weight (g)",
       y = "Count",
       title = "Juvenile Snowshoe Hare Count by Weight (CORNELIUS)") +
  theme(plot.title = element_text(hjust = 0.5))
```

```{r}
# Make exploratory qq plot to compare sex effect on weight
ggplot(data = male_female, aes(sample = weight)) +
  geom_qq() +
  facet_wrap(~sex) +
  labs(x = "Theoretical",
       y = "Sample",
       title = "Sex Effect on Weight (CORNELIUS)") +
  theme(plot.title = element_text(hjust = 0.5))
```

#### D. Relationship between juvenile weight & hindfoot length

```{r, r juvenile_model}
# Do linear regression
juvenile_lm <- lm(weight ~ hindft, data = hares)

# Get tidy versions of the model output to call later on in text
juvenile_lm_tidy <- tidy(juvenile_lm)
juvenile_lm_glance <- glance(juvenile_lm)

# Metrics at a glance: 
juvenile_lm_out <- broom::glance(juvenile_lm)
juvenile_lm_out

# Get Pearson's r correlation
juvenile_cor <- cor.test(hares$hindft, hares$weight)

# Tidy version of correlation output:
juvenile_cor_tidy <- tidy(juvenile_cor)

juvenile_lm_tidy

juvenile_cor_tidy
```
```{r}
plot(juvenile_lm)
```

```{r, include = FALSE}
# Get the intercept: 
juvenile_int <- juvenile_lm_tidy$estimate[1]
juvenile_int

# Then to get the hindfoot coefficient:
juvenile_coef <- juvenile_lm_tidy$estimate[2]
juvenile_coef
```


```{r}
# Use the `equatiomatic` package to get the output as a LaTeX equation
extract_eq(model = juvenile_lm, use_coefs = TRUE)
```

```{r}
# Make subset of juvenile hindfoot length and weight
weight_length <- hares %>%
  filter(age == "j") %>%
  select(hindft, weight)
  
# Make ggplot to visualize relationship between juvenile hindfoot length and weight
ggplot(data = weight_length, aes(x = hindft, y = weight)) +
  geom_point(color = "red") +
  labs(x = "Hare Hindfoot Length (mm)",
       y = "Hare Weight (g)") +
  theme(plot.title = element_text(hjust = 0.5)) +
  theme_pubr() +
  geom_smooth(method = "lm", se = FALSE, color = "gray30") +
  labs(x = "Hindfoot length (mm)",
       y = "Weight (g)") +
  theme(legend.position = c(0.15, 0.7))
```

**Figure 3.** Relationship between hindfoot length (mm) and weight (g) for juvenile snowshoe hares. Linear model summary: $\beta$~1~ = `r round(juvenile_lm_tidy$estimate[2], 2)` g mm^-1^, p < 0.001, R^2^ = `r round(juvenile_lm_glance$r.squared,2)`, Pearson's *r* = `r round(juvenile_cor_tidy$estimate,2)`). Data: Kielland et al. (2017).

The relationship between juvenile snowshoe hare hindfoot length and weight appears relatively linear (Figure 3). Simple linear regression revealed that hindfoot length loosely predicts snowshoe hare weight (p < 0.001, R^2^ = `r round(juvenile_lm_glance$r.squared,2)`) with an average slope of $\beta$ = `r round(juvenile_lm_tidy$estimate[2], 2)` g mm^-1^ (i.e., for each one millimeter increase in snowshoe hindfoot length we expect an average increase in weight of `r round(juvenile_lm_tidy$estimate[2], 2)` g). Hindfoot length and weight have medium correlation (Pearson's *r* = `r round(juvenile_cor_tidy$estimate,2)`, p < 0.001). Diagnostic plots (not included) reveal non-normally distributed and heteroscedastic residuals??

## Summary

Exploratory data analysis reveals the following initial findings:
- Male juvenile snowshoe hares weigh more than mean flipper length than female juvenile snowshoe hares (the difference is significant, and the effect size is small)
- Hindfoot length appears linearly related with weight across males and females; the relationship (by simple linear regression) is significant, with a slope (expected average increase in weight with a 1 mm increase in flipper length) of `r round(penguin_lm_tidy$estimate[2], 2)` g mm^-1^.

Next steps: 
- Compare hindfoots statistically (one-way ANOVA), including hare sex as a covariate
- Multiple linear regression to understand the relationship between age, sex, hindfoot length and weight
- Analysis of grids to explore size correlation

## Citations

Kielland, K., F.S. Chapin, R.W. Ruess, and Bonanza Creek LTER. 2017. Snowshoe hare physical data in Bonanza Creek Experimental Forest: 1999-Present ver 22. Environmental Data Initiative. (https://doi.org/10.6073/pasta/03dce4856d79b91557d8e6ce2cbcdc14)

R Core Team (2021). R: A language and environment for statistical computing. R Foundation for
Statistical Computing, Vienna, Austria. URL https://www.R-project.org/.
  
RStudio Team (2021). RStudio: Integrated Development Environment for R. RStudio, PBC, Boston, MA
URL: (http://www.rstudio.com/)
