---
title: "Task 2: Snowshoe Report (CORNELIUS)"
author: "Andrea Cornelius"
date: "11/3/2021"
output: html_document
---
*Introduction:* A 4 - 5 sentence "mini-introduction" to what this study is exploring, including basic background information about the subject, site and study (with citations as needed), and a brief overview of what is included in the exploratory report. You should probably do this last so you know what’s contained in the report.

*Data and analyses:* Kielland, K., F.S. Chapin, R.W. Ruess, and Bonanza Creek LTER. 2017. Snowshoe hare physical data in Bonanza Creek Experimental Forest: 1999-Present ver 22. Environmental Data Initiative. (https://doi.org/10.6073/pasta/03dce4856d79b91557d8e6ce2cbcdc14) (Accessed 2021-11-07).
A 3 - 4 sentence section in which you briefly describe the data, including citation & credit as needed, and provide an overview of the main analysis, statistical methods and tools used.

# Install packages
```{r setup, include=FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
# Attach packages
library(here)
library(janitor)
library(tidyverse)
library(lubridate)
library(kableExtra)
```

# Read in the data
```{r}
# Read in data
hares <- read_csv(here("data", "bonanza_hares.csv"))
```
# Data wrangling: date classification
```{r}
# Date classification
date <- as.Date("1998-11-26") # Classify date

date

class(date) # Check the class
```

# Juvenile hare initial investigation

## Annual juvenile hare trap counts
 
# Data wrangling to Count the total number of juvenile hare trappings during each year of the study
```{r}
# Count the total number of juvenile hare trappings during each year of the study
hare_subset <- hares %>% # Pull from hares data
  mutate(date_new = mdy(date)) %>% # Convert class to date
  mutate(year = year(date_new)) %>% # Add new column for year
  filter(age == "j") %>% # Filter age to juvenile
  group_by(year) %>% # Group by year
  summarize(count = n(),
            mean = mean(count),
            median = median(count)) # Summarize count, mean, median
```

# Include a figure caption below the figure in your report. 

# Data visualization of the total number of juvenile hare trappings during each year of the study
```{r}
ggplot(data = hare_subset, aes(x = year, y = count)) +
  geom_bar(stat = "identity", fill="#f68060") +
  scale_x_continuous(limits=c(1998, 2013), breaks = c(1999, 2000, 2001, 2002, 2003, 2004, 2005, 2006, 2007, 2008, 2009, 2010, 2011, 2012)) +
  scale_y_continuous(limits=c(0, 130), breaks = c(0, 25, 50, 75, 100, 125)) +
  theme_gray() +
   labs(x = "Year",
       y = "Juvenile Snowshoe Hare Trappings", 
       title = "Juvenile Snowshoe Hares Trapped at Bonanza Creek LTER (CORNELIUS)") + 
  theme(plot.title = element_text(hjust = 0.5))
```

**Figure 1: The snowshoe hare population was largest in 1999, at **

## Juvenile hare weights

# Data wrangling to count the total number of juvenile hare trappings by grid and sex
```{r}
hare_subset_2 <- hares %>% # Pull from hares data
  filter(age == "j") %>% # Filter age to juvenile
  group_by(grid, sex) %>%
  select(-date, -time, -trap, -l_ear, -r_ear, -hindft, -notes, -b_key, -session_id, -study)

hare_subset_2$grid <- gsub("bonrip", "Bonanza Riparian", hare_subset_2$grid) 
hare_subset_2$grid <- gsub("bonmat", "Bonanza Mature", hare_subset_2$grid)
hare_subset_2$grid <- gsub("bonbs", "Bonanza Black Spruce", hare_subset_2$grid)
hare_subset_2$sex <- gsub("m", "Male", hare_subset_2$sex)
hare_subset_2$sex <- gsub("f", "Female", hare_subset_2$sex)
hare_subset_2$sex <- gsub("NA", "Unknown", hare_subset_2$sex)
``` 

# Data visualization of the total number of juvenile hare trappings by grid and sex
```{r}
ggplot(data = hare_subset_2, aes(x = sex, y = weight, color = sex)) +
  geom_point(size = 2) + 
  scale_color_manual(values = c("pink", "blue", "grey"), 
                     name = "Penguin species") +
  facet_wrap(~ grid) +
  labs(x = "Sex",
       y = "Juvenile Snowshoe Hare Weight (g)",
       title = "Juvenile Snowshoe Hare Weights by Sex and Site (CORNELIUS)") +
  theme_bw() +
  theme(legend.position = "none",
        plot.title = element_text(hjust = 0.5),
        axis.title.x = element_text(size = 10))
```

**Figure 2:**
