---
title: "Task 2: Snowshoe Report (CORNELIUS)"
author: "Andrea Cornelius"
date: "11/3/2021"
output: html_document
---
*Introduction:* A 4 - 5 sentence "mini-introduction" to what this study is exploring, including basic background information about the subject, site and study (with citations as needed), and a brief overview of what is included in the exploratory report. You should probably do this last so you know what’s contained in the report.

*Data and analyses:* A 3 - 4 sentence section in which you briefly describe the data, including citation & credit as needed, and provide an overview of the main analysis, statistical methods and tools used.

# Install packages
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Attach packages
library(here)
library(janitor)
library(tidyverse)
library(lubridate)
library(kableExtra)
```

# Read in the data
```{r}
# Read in data
hares <- read_csv(here("data", "bonanza_hares.csv"))
```
# Data wrangling: date classification
```{r}
# Date classification
date <- as.Date("1998-11-26") # Classify date

date

class(date) # Check the class
```

# Juvenile hare initial investigation

## Annual juvenile hare trap counts
 
# Data wrangling to Count the total number of juvenile hare trappings during each year of the study
```{r}
# Count the total number of juvenile hare trappings during each year of the study
hare_subset <- hares %>% # Pull from hares data
  mutate(date_new = mdy(date)) %>% # Convert class to date
  mutate(year = year(date_new)) %>% # Add new column for year
  filter(age == "j") %>%
  group_by(year) %>%
  summarize(count = n(),
            mean = mean(count),
            median = median(count))
```

# Include a figure caption below the figure in your report. 

```{r}
ggplot(data = hare_subset, aes(x = year, y = juvenile_trappings)) +
  geom_bar(stat = "identity", fill="#f68060") +
  scale_x_continuous(limits=c(1998, 2013), breaks = c(1999, 2000, 2001, 2002, 2003, 2004, 2005, 2006, 2007, 2008, 2009, 2010, 2011, 2012)) +
  scale_y_continuous(limits=c(0, 130), breaks = c(0, 25, 50, 75, 100, 125)) +
  theme_gray() +
   labs(x = "Year",
       y = "Juvenile Snowshoe Hare Trappings", 
       title = "Juvenile Snowshoe Hare Trappings at Bonanza Creek LTER, 1999-2012 (CORNELIUS)")
```

## Visualize juvenile hare weights

```{r}
```


```{r}
hare_subset_2 <- hares %>% # Pull from hares data
  filter(age == "j") 
  
```

