---
title: "Task 2: Snowshoe Report (CORNELIUS)"
author: "Andrea Cornelius"
date: "11/3/2021"
output: html_document
---
*Introduction:* A 4 - 5 sentence "mini-introduction" to what this study is exploring, including basic background information about the subject, site and study (with citations as needed), and a brief overview of what is included in the exploratory report. You should probably do this last so you know what’s contained in the report.

*Data and analyses:* A 3 - 4 sentence section in which you briefly describe the data, including citation & credit as needed, and provide an overview of the main analysis, statistical methods and tools used.

# Install packages
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Attach packages
library(here)
library(janitor)
library(tidyverse)
library(lubridate)
library(kableExtra)
```

# Read in the data
```{r}
# Read in data
hares <- read_csv(here("data", "bonanza_hares.csv"))
```
# Data wrangling: date classification
```{r}
# Date classification
date <- as.Date("1998-11-26") # Classify date

date

class(date) # Check the class
```

# Count the total number of juvenile hare trappings during each year of the study, and create a finalized data visualization of the counts by year. Include a figure caption below the figure in your report. 
```{r}
hare_dates <- hares %>%
  mutate(date_new = mdy(date)) %>% # Convert class to date
  mutate(year = year(date_new))
```

